---
title: "Project 3"
author: "Joy Zhou"
date: "2023-10-31"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The `Behavioral Risk Factor Surveillance System (BRFSS)` is an annual health-related telephone survey conducted by the Centers for Disease Control and Prevention (CDC) since 1984. It is designed to gather standardized, state-specific information on preventive health practices and risk behaviors associated with chronic diseases, injuries, and preventable infectious diseases within the adult population.

Diabetes stands as one of the most prevalent chronic diseases in the United States, affecting millions of Americans annually and imposing a substantial economic burden on the nation. 

For this project, we will use the `diabetes _ binary _ health _ indicators _ BRFSS2015.csv` dataset, sourced from [Diabetes Health Indicators Dataset](https://www.kaggle.com/datasets/alexteboul/diabetes-health-indicators-dataset/).

This dataset comprises 253,680 survey responses to the CDC's BRFSS2015. The target variable `Diabetes_binary` has two classes: 0 for no diabetes, and 1 for prediabetes or diabetes. This dataset encompasses 21 feature variables and is not balanced, with Diabetes_binary serving as the response variable. The information of variable can be found [here](https://www.kaggle.com/datasets/alexteboul/diabetes-health-indicators-dataset/?select=diabetes_binary_health_indicators_BRFSS2015.csv)  

+ Diabetes_binary: 0 = no diabetes 1 = prediabetes or diabetes  
+ HighBP: High blood pressure  
+ HighChol: High cholesterol   
+ CholCheck: 0 = no cholesterol check in 5 years 1 = yes cholesterol check in 5 years  
+ BMI: Body Mass Index   
+ Smoker: Have you smoked at least 100 cigarettes in your entire life? [Note: 5 packs = 100 cigarettes] 0 = no 1 = yes  
+ Stroke: 0 = no 1 = yes  
+ HeartDiseaseorAttack: coronary heart disease (CHD) or myocardial infarction (MI) 0 = no 1 = yes  
+ PhysActivity: physical activity in past 30 days - not including job 0 = no 1 = yes   
+ Fruits: Consume Fruit 1 or more times per day 0 = no 1 = yes  
+ Veggies: Consume Vegetables 1 or more times per day 0 = no 1 = yes  
+ HvyAlcoholConsump: (adult men >=14 drinks per week and adult women>=7 drinks per week) 0 = no 1 = yes  
+ AnyHealthcare: Health care coverage 0 = no 1 = yes  
+ NoDocbcCost: Was there a time in the past 12 months when you needed to see a doctor but could not because of cost? 0 = no 1 = yes  
+ GenHlth: in general your health is: scale 1-5 1 = excellent 2 = very good 3 = good 4 = fair 5 = poor  
+ MentHlth: days of poor mental health scale 1-30 days  
+ PhysHlth: physical illness or injury days in past 30 days scale 1-30  
+ DiffWalk: Do you have serious difficulty walking or climbing stairs? 0 = no 1 = yes  
+ Sex: 0 = female 1 = male  
+ Age: 13-level age category 1 = 18-24 2 = 25-29 3 = 30-34 4 = 35-39 5 = 40-44 6 = 45-49 7 = 50-54 8 = 55-59 9 = 60-64 10 = 65-69 11 = 70-74 12 = 75-79 13 = 80 or older  
+ Education: scale 1-6 1 = Never attended school or only kindergarten 2 = Grades 1 through 8 (Elementary) 3 = Grades 9 through 11 (Some high school) 4 = Grade 12 or GED (High school graduate) 5 = College 1 year to 3 years (Some college or technical school) 6 = College 4 years or more (College graduate)  
+ Income: scale 1-8 1 = less than $10,000 5 = less than $35,000 8 = $75,000 or more  

# Data
## Read in data
```{r, message=FALSE}
library(dplyr)
library(readr)
diabetes <- as_tibble(read.csv("diabetes_binary_health_indicators_BRFSS2015.csv", header = TRUE))
head(diabetes)
Diabetes_binary <- as.factor(diabetes$Diabetes_binary)
Income <- factor(diabetes$Income, levels = c("1", "2", "3", "4", "5", "6", "7", "8"))
```

## Subsetting the dataset based on education level

```{r }
# Never attended school or only kindergarten to Grades 1 through 8 (Elementary)
diabetes1_2 <- diabetes %>% filter(Education %in% c(1, 2))
# Grades 9 through 11 (Some high school) 
diabetes3 <- diabetes %>% filter(Education == 3)
# Grade 12 or GED (High school graduate) 
diabetes4 <- diabetes %>% filter(Education == 4)
# College 1 year to 3 years (Some college or technical school) 
diabetes5 <- diabetes %>% filter(Education == 5)
# College 4 years or more (College graduate) 
diabetes6 <- diabetes %>% filter(Education == 6)
```

## EDA for diabetes3
```{r}
# Checking missing values of columns in diabetes3
missing_values <- colSums(is.na(diabetes3))
missing_values # no missing values
# Describing 
library(knitr)
kable(summary(diabetes3))
```
result shows that there is no missing values in dataset diabetes3

```{r}
# Set the figure size (in inches)
options(repr.plot.width = 8, repr.plot.height = 6)




```

